 

### 2022-05-17 scripts/review.Rmd
Create additional data summary

>>>>

#### Create data of all sampled studies

```{r}
load("../data/screen/sample.Rda")
abstract <- read.csv("../data/screen/abstract_screening.csv")
fulltext <- read.csv("../data/screen/fulltext_screening.csv")
fullsample <- sampled_data[, c("sid", "title", "authors", "year", "doi")]
get_status <- function(sid) {
  if (!(sid %in% fulltext$sid)) {
    status <- "excluded_abstract"
    reason <- abstract$decision[sid]
  } else {
    if (is.na(fulltext$decision[sid])) {
      status <- "included"
      reason <- NA
    } else {
      status <- "excluded_fulltext"
      reason <- fulltext$decision[sid]
    }
  }
  data.frame(status = status, reason = reason)
}
fullsample <- cbind(
  fullsample, 
  purrr::map_dfr(seq_len(nrow(fullsample)), get_status)
)

fullsample$reason <- factor(
  fullsample$reason, 
  levels = c("t", "l", "r", "h", "m", "c", "e", "d", "o"),
  labels = c("report not retrieved", "not in English", "no primary research", 
             "reasearch not in humans", "VO2 only estimated", "no appropriate test protocol (VO2 not continously measured)", 
             "no appropriate test protocol (no exhaustion)", "no appropriate test protocol (longer than 20 minutes)", "no information regarding the exclusion criteria")
)

write.csv(fullsample, file = "../data/fullsample.csv", na = "", row.names = FALSE)
```

### 2022-05-17 thesis/methods.qmd
Create additional data summary

I conducted all analyses using R Version 4.2.0 [@rcoreteam2022] within R Studio Version 2022.2.2.485 [@rstudioteam2022]. The thesis was entirely written in Quarto Version 0.9.380 [@allaire2022a]. The attached packages and default settings are documented in @sec-tech.

@@

To determine the influence the most common data processing strategies have on the determination of V̇O~2max~, I compared them on a set of already collected gas exchange data from ramp tests in running.

@@

A total of N = 72 exercise tests were analysed for this study. Due to a miscalculation, the preregistration had incorrectly stated a number of 76 tests. The data was from previous research on the metabolic profile of endurance runners [@quittmann_unpubl_a; @quittmann_unpubl_b]. The tested individuals were experienced distance runners (15 female, 54 male; three of the males participated in both studies). The V̇O~2max~ tests were conducted in March to September 2019 [@quittmann_unpubl_a] and March to October 2021 [@quittmann_unpubl_b] using an identical exercise protocol. Participants run on a treadmill (saturn 300/100, h/p/cosmos sports & medical 127 GmbH, Nussdorf-Traunstein, Germany) with 1% inclination for ten minutes at a velocity of 2.8 m·s^-1^ as a warm-up. After preparing the gas exchange measures, they started a ramp protocol with an initial speed of 2.8 m·s^-1^ for two minutes and subsequently increased velocity by 0.15 m·s^-1^ every 30 seconds. The researchers provided verbal encouragement and terminated the exercise when the participants reached subjective exhaustion.

@@

Moving time based averages were calculated by first linearly interpolating the breath-by-breath data to seconds. Subsequently a (center aligned) moving average was calculated over a defined timespan. These processing steps are implemented in the spiro package [@nolte2022]. For calculating the moving average over 30 seconds for example, I used the functions `spiro(data) |> spiro_max(30)`.

Binned time averages were calculated using a custom function (available in the [analysis script](https://github.com/smnnlt/vo2max-processing/blob/master/scripts/comparison.md) on GitHub). Breath-by-breath data was initially interpolated to full seconds and then binned into consecutive interval of a constant length. The average of each interval was aligned to its center. Incomplete intervals (i.e. the last seconds of measurement) were not considered in the analysis. Note that some authors use a different procedure for determining their bins, starting by the end point of the measurement. However, defining bins beginning by the start of the measurement is a common output option for many gas exchange data analysis software (e.g. Cosmed Omnia).

@@

Butterworth filters are a class of recursive digital filters. @robergs2010 and @weir2004 argue that these more advanced processing strategies are superior to the traditional moving or binned average approach for analysing gas exchange data. However @robergs2010 missed to account for the time lag introduced by Butterworth filters. Therefore I applied a zero-phase Butterworth filter by means of a forward-backward filtering. While this effectively zeroes out the time lag, the resulting filter is non-causal, which means it cannot be used online (i.e. in real time). However, for the present application, an offline filter is sufficient. The forward-backward filtering also introduces transients at both ends of the signal [@gustafsson1996]. I therefore padded the reverted signal at both the start and the end of the array (identical to the 'even padtype' in Python's `scipy.signal.filtfilt()` function [@virtanen2020]). I used 3 as the order parameter and 0.04 as the cut-off frequency for each filter [@robergs2010]. Note that due to the forward-backward approach both the overall order and overall cut-off frequency are different from these values. The described approach is implemented in the spiro package [@nolte2022], and can be used by calling, for example, the functions `spiro(data) |> spiro_max("0.04fz3")`.

>>>>

I conducted all analyses using R Version 4.2.0 [@rcoreteam2022] in the R Studio IDE Version 2022.2.2.485 [@rstudioteam2022]. The thesis was entirely written in Quarto Version 0.9.380 [@allaire2022a]. The attached packages and default settings are documented in @sec-tech.

@@

To determine the influence of the most common data processing strategies on the determination of V̇O~2max~, I compared them on a set of already collected gas exchange data from ramp tests in running.

@@

A total of N = 72 exercise tests were analysed for this study. Due to a miscalculation, the preregistration had incorrectly stated a number of 76 tests. The data were from previous research on the metabolic profile of endurance runners [@quittmann_unpubl_a; @quittmann_unpubl_b]. The tested individuals were experienced distance runners (15 female, 54 male; three of the males participated in both studies). The V̇O~2max~ tests were conducted in March to September 2019 [@quittmann_unpubl_a] and March to October 2021 [@quittmann_unpubl_b] using an identical exercise protocol. Participants run on a treadmill (saturn 300/100, h/p/cosmos sports & medical 127 GmbH, Nussdorf-Traunstein, Germany) with 1% inclination for ten minutes at a velocity of 2.8 m·s^-1^ as a warm-up. After preparing the gas exchange measures, they started a ramp protocol with an initial speed of 2.8 m·s^-1^ for two minutes and subsequently increased velocity by 0.15 m·s^-1^ every 30 seconds. The researchers provided verbal encouragement and terminated the exercise when the participants reached subjective exhaustion.

@@

Moving time-based averages were calculated by first linearly interpolating the breath-by-breath data to seconds. Subsequently a (center aligned) moving average was calculated over a defined timespan. These processing steps are implemented in the spiro package [@nolte2022]. For calculating the moving average over 30 seconds for example, I used the functions `spiro(data) |> spiro_max(30)`.

Binned time averages were calculated using a custom function (available in the [analysis script](https://github.com/smnnlt/vo2max-processing/blob/master/scripts/comparison.md) on GitHub). Breath-by-breath data were initially interpolated to full seconds and then binned into consecutive intervals of constant lengths. The average of each interval was aligned to its center. Incomplete intervals (i.e. the last seconds of measurement) were not considered in the analysis. Note that some authors use a different procedure for determining their bins, starting by the end point of the measurement. However, defining bins beginning by the start of the measurement is a common output option for many gas exchange data analysis software (e.g. Cosmed Omnia).

@@

Butterworth filters are a class of recursive digital filters. Robergs et al. @robergs2010 and Weir et al. @weir2004 argue that these more advanced processing strategies are superior to the traditional moving or binned average approach for analysing gas exchange data. However Robergs et al. @robergs2010 missed to account for the time lag introduced by Butterworth filters. Therefore I applied a zero-phase Butterworth filter by means of forward-backward filtering. While this effectively zeroes out the time lag, the resulting filter is non-causal, which means it cannot be used online (i.e. in real time). However, for the present application, an offline filter is sufficient. The forward-backward filtering also introduces transients at both ends of the signal [@gustafsson1996]. I therefore padded the reverted signal at both the start and the end of the array (identical to the 'even padtype' in Python's `scipy.signal.filtfilt()` function [@virtanen2020]). I used 3 as the order parameter and 0.04 as the low-pass cut-off frequency for each filter [@robergs2010]. Note that due to the forward-backward approach both the filter's overall order and overall cut-off frequency are different from these values. The described approach is implemented in the spiro package [@nolte2022], and can be used by calling, for example, the functions `spiro(data) |> spiro_max("0.04fz3")`.
